---
title: "ECON104_Project"
author: 'Group Members: Alden Do, Ulisses Hernandez, Fabian Diaz, and Sai Rachakonda'
date: "2026-03-05"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

set.seed(123)

# Core packages
library(tidyverse)
library(janitor)

# Variable selection
library(Boruta)

# Regression + diagnostics
library(car)
library(lmtest)
library(sandwich)

# Summary tables
library(modelsummary)

# Normality test
library(tseries)

# Bootstrap + CV
library(boot)
library(caret)
```

```{r data_import_cleaning}
# 1) Import raw data
# Keep the CSV in the same folder as this .Rmd file

df_raw <- read.csv("Data/Raw/Amazon Customer Behavior Survey 3.csv", stringsAsFactors = FALSE)

# 2) Standardize variable names and trim extra spaces in text cells
df_clean <- df_raw %>%
  clean_names() %>%
  mutate(across(where(is.character), stringr::str_squish))

# 3) Convert categorical variables to factors
# Treat all character columns as categorical for this survey dataset
df_clean <- df_clean %>%
  mutate(across(where(is.character), as.factor))

# 4) Flag missing values (do not remove yet)
# Convert blank strings to NA first so they are counted properly

df_clean <- df_clean %>%
  mutate(across(where(is.factor), ~ forcats::fct_explicit_na(., na_level = "Missing")))

missing_summary <- tibble(
  variable = names(df_clean),
  missing_count = sapply(df_clean, function(x) sum(is.na(x) | x == "Missing")),
  missing_percent = round(sapply(df_clean, function(x) mean(is.na(x) | x == "Missing")) * 100, 2)
)

# 5) Flag numeric outliers using the IQR rule
# Do NOT delete observations yet; just create indicator columns

numeric_vars <- df_clean %>%
  select(where(is.numeric)) %>%
  names()

for (v in numeric_vars) {
  q1 <- quantile(df_clean[[v]], 0.25, na.rm = TRUE)
  q3 <- quantile(df_clean[[v]], 0.75, na.rm = TRUE)
  iqr_value <- q3 - q1
  lower_bound <- q1 - 1.5 * iqr_value
  upper_bound <- q3 + 1.5 * iqr_value

  df_clean[[paste0(v, "_outlier_flag")]] <- ifelse(
    is.na(df_clean[[v]]),
    NA,
    df_clean[[v]] < lower_bound | df_clean[[v]] > upper_bound
  )
}

# 6) Quick checks

glimpse(df_clean)
dim(df_clean)
colnames(df_clean)
missing_summary
```

## Variables Description

## Variables Description

| Variable Name | Definition (in words) | Units | Type | Expected Sign |
|---|---|---|---|---|
| purchase_frequency | How often a respondent makes purchases on Amazon | Ordered frequency category | Factor | Dependent variable |
| purchase_categories | Product categories most frequently purchased | Category | Factor | ? |
| browsing_frequency | How often a respondent browses Amazon | Ordered frequency category | Factor | + |
| customer_reviews_importance | Importance of online reviews in purchase decisions | Rating scale | Numeric | + |
| add_to_cart_browsing | Whether the respondent adds items to cart while browsing | Category | Factor | + |
| saveforlater_frequency | How often the respondent uses the save-for-later feature | Ordered frequency category | Factor | + |
| recommendation_helpfulness | Whether personalized recommendations are considered helpful | Category | Factor | + |
| rating_accuracy | Perceived accuracy of product ratings | Rating scale | Numeric | + |
| shopping_satisfaction | Overall satisfaction with Amazon shopping experience | Rating scale | Numeric | + |
| age | Respondent age | Years | Numeric | ? |
| gender | Respondent gender | Category | Factor | ? |

#1. Economic Theory and Hypothesis 
## (a) Economic Research Question: 
Research Question: What platform features and consumer behaviors influence how frequently consumers make purchases on Amazon?

## (b) Economic Theory Review:
The analysis reveals that the strategic use of customer reviews, advanced product search filters, and features like “save for later” markedly enhance the likelihood of frequent purchases. Global e-commerce has sustained strong growth since the pandemic and is projected to exceed $8 trillion by 2027, indicating a structural shift in consumer purchasing toward digital channels (Uzunoglu, 2024; Lin, 2024). Purchase frequency, defined as the number of transactions per customer within a given timeframe, is a more direct measure of engagement and loyalty than one off conversion (Hohenberg & Taylor, 2020). Online reviews shape purchase intentions and decision confidence (Ahn & Lee, 2024; Chen et al., 2022; Zhang et al., 2018), and that search and filtering facilitate discovery (Deng et al., 2021). Secondly, researchers often examine predictors in isolation, which obscures the joint influence of category cadence and interface use on behaviour. Thirdly, there is a limited operationalization of Amazon's specific affordances, such as the "save for later" feature, review reliability vs. importance, and advanced filters, as potential determinants of repeat purchasing product category, review perceptions (importance and reliability), search method (including filters), “save for later” usage, and browsing frequency relate to the likelihood of purchasing on Amazon a few times per month.Category effects dominate: groceries and clothing/fashion correspond to the highest likelihood of purchasing a few times per month
Individuals who browse the platform a few times per week are more likely to make purchases on a monthly basis than the sample average.Use of the "save for later" functions are each associated with a higher frequency

Consumer decision-making in online marketplaces can be analyzed using information asymmetry theory and consumer search theory. In online markets like Amazon, buyers cannot physically inspect products before purchase, creating uncertainty about quality. According to Akerlof’s (1970) “Market for Lemons” framework, when information asymmetry exists between buyers and sellers, market efficiency may decline unless credible signals reduce uncertainty. Online reviews function as a signaling mechanism that reduces information asymmetry by providing information about product quality from previous buyers. This increases trust and lowers perceived risk, which can increase the probability and frequency of purchase.

Additionally, consumer search theory (Stigler, 1961) suggests that individuals gather information as long as the marginal benefit of additional information exceeds the marginal cost. On platforms like Amazon, reviews significantly reduce search costs by aggregating product experiences in one place. Lower search costs and increased trust can shift consumer demand outward, increasing purchase frequency. Empirical research also supports this mechanism. For example, Chevalier and Mayzlin (2006) find that positive online reviews significantly increase sales, demonstrating that user-generated content directly influences consumer demand in digital markets.

Source 1: 
Source 2: 

## (c) Testable Economic Hypothesis
Based on consumer demand theory, individuals with higher income are expected to spend more on goods and services. Therefore, we hypothesize that income will have a positive relationship with consumer purchasing behavior. In other words, as income increases, the likelihood or amount of consumer purchases should increase. Additionally, greater exposure to advertising is expected to increase consumer awareness and influence purchasing decisions. Therefore, advertising exposure should have a positive effect on purchase behavior.

## (d) Expected Signs of Coefficients:
Explain the expected direction (+ or -) of key variables based on economic theory.

# 2. Variable Selection 

## (a) Boruta Algorithm (Quantitative Predictors)
Explain the Boruta method and include code. 
```{r boruta}
boruta_model <- Boruta(purchase_frequency ~ ., data = df_clean)

print(boruta_model)
plot(boruta_model)
```

## (b) Group Comparison Test
```{r group-comparison}
# t-test or ANOVA code here
```

## (c) Final Predictor Set 
List the final predictors chosen for the rest of the analysis. 

# 3. Descripitve Analysis and Exploratory Data Analysis 
## (a) Dataset Description 
```{r summary}
summary(df_clean)
```
Describe: dataset source, time period, sample, variable definitions

## (b) Exploratory Visualizations 
# histograms
# scatterplots
# correlation plots

## (c) Descripitve Statistics Table 
# summary statistics 

## (d) Density Plots 
# density plots 

## (e) Nonlinearity and Transformations 
Discuss transformations such as log or polynomial terms 

## (f) Outliers 
Explain any outliers and how you handled them 

## (g) Missing Data 
Explain NA removal or imputation 

# 4. Model Building and Selection 
Model 1 
#regression model 

Model 2 
#regreession model

Model 3 
#regreession model 

Multicollinearity (VIF)
#VIF test 

Model Misspecification (RESET test)
#RESET test 

Diagnostic Plots 
#residual plots 

Model Comparison 
# AIC / BIC
Explain:

#5. Model Diagnostics and Robustness 
Linearity: Residual v. Fitted Plots 

Homoskedasticity 
#Breusch-Pagan test 

Normality 
# QQ plot
# Jarque-Bera test

Outliers and Influence
# Cook's distance

Bootstrap Analysis
# bootstrap code

Cross Validation
# 5-fold or 10-fold CV

# 6. Results and Economic Interpretation
# regression table
Explain: coefficient signs,economic meaning, statistical significance, confidence intervals, R²

# 7. Limitations and Extensions
data limitations, potential bias, possible future research

# 8. Code Quality and Reproducibility
Explain: reproducibility, seed setting, well-commented code

