---
title: "ECON104_Project"
author: "Group Members: Alden Do, Ulisses Hernandez, Fabian Diaz, and Sai Rachakonda" 
date: "2026-03-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

set.seed(123)

#Core Packages 
library(tidyverse)
library(janitor)
#Variable Selection
library(Boruta)
#Regression + diagnostics 
library(car)
library(lmtest)
library(sandwich)
#Summary tables 
library(modelsummary)
#Normality test 
library(tseries)
# Bootstrap + CV
library(boot)
library(caret)

#Loading Data 
df_raw <- read.csv("Amazon Customer Behavior Survey 3.csv", stringsAsFactors = FALSE)

#Cleaning Column Names + Trimming Extra Spaces 
df <- df_raw %>%
  clean_names() %>%
  mutate(across(where(is.character), stringr::str_squish))

#Quick Chekcs 
glimpse(df)
dim(df)
colnames(df) 

```

#1. Economic Theory and Hypothesis 
## (a) Economic Research Question: 
Research Question: What platform features and consumer behaviors influence how frequently consumers make purchases on Amazon?

## (b) Economic Theory Review:
The analysis reveals that the strategic use of customer reviews, advanced product search filters, and features like “save for later” markedly enhance the likelihood of frequent purchases. Global e-commerce has sustained strong growth since the pandemic and is projected to exceed $8 trillion by 2027, indicating a structural shift in consumer purchasing toward digital channels (Uzunoglu, 2024; Lin, 2024). Purchase frequency, defined as the number of transactions per customer within a given timeframe, is a more direct measure of engagement and loyalty than one off conversion (Hohenberg & Taylor, 2020). Online reviews shape purchase intentions and decision confidence (Ahn & Lee, 2024; Chen et al., 2022; Zhang et al., 2018), and that search and filtering facilitate discovery (Deng et al., 2021). Secondly, researchers often examine predictors in isolation, which obscures the joint influence of category cadence and interface use on behaviour. Thirdly, there is a limited operationalization of Amazon's specific affordances, such as the "save for later" feature, review reliability vs. importance, and advanced filters, as potential determinants of repeat purchasing product category, review perceptions (importance and reliability), search method (including filters), “save for later” usage, and browsing frequency relate to the likelihood of purchasing on Amazon a few times per month.Category effects dominate: groceries and clothing/fashion correspond to the highest likelihood of purchasing a few times per month
Individuals who browse the platform a few times per week are more likely to make purchases on a monthly basis than the sample average.Use of the "save for later" functions are each associated with a higher frequency

Consumer decision-making in online marketplaces can be analyzed using information asymmetry theory and consumer search theory. In online markets like Amazon, buyers cannot physically inspect products before purchase, creating uncertainty about quality. According to Akerlof’s (1970) “Market for Lemons” framework, when information asymmetry exists between buyers and sellers, market efficiency may decline unless credible signals reduce uncertainty. Online reviews function as a signaling mechanism that reduces information asymmetry by providing information about product quality from previous buyers. This increases trust and lowers perceived risk, which can increase the probability and frequency of purchase.

Additionally, consumer search theory (Stigler, 1961) suggests that individuals gather information as long as the marginal benefit of additional information exceeds the marginal cost. On platforms like Amazon, reviews significantly reduce search costs by aggregating product experiences in one place. Lower search costs and increased trust can shift consumer demand outward, increasing purchase frequency. Empirical research also supports this mechanism. For example, Chevalier and Mayzlin (2006) find that positive online reviews significantly increase sales, demonstrating that user-generated content directly influences consumer demand in digital markets.

Source 1: 
Source 2: 

## (c) Testable Economic Hypothesis
Based on consumer demand theory, individuals with higher income are expected to spend more on goods and services. Therefore, we hypothesize that income will have a positive relationship with consumer purchasing behavior. In other words, as income increases, the likelihood or amount of consumer purchases should increase. Additionally, greater exposure to advertising is expected to increase consumer awareness and influence purchasing decisions. Therefore, advertising exposure should have a positive effect on purchase behavior.

## (d) Expected Signs of Coefficients:

**Perceived importance and reliability of customer reviews (+):**  
When shopping online, such as on Amazon, consumers cannot directly examine products before purchasing. This creates uncertainty about product quality. Customer reviews help reduce that uncertainty by giving information from previous buyers and increasing consumer confidence. Consumers who see reviews as important and trustworthy are more likely to feel comfortable making purchases, so the expected sign is positive.

**Use of search filters and advanced search tools (+):**  
Using search filters and advanced search tools should have a positive relationship with purchase frequency. These tools make it easier for consumers to narrow down the large number of products on Amazon and quickly find items that match their preferences. By reducing the time and effort required to search, these features lower search costs and make purchases more likely.

**“Save for Later” feature (+):**  
The “Save for Later” feature is also expected to have a positive coefficient. This feature allows consumers to keep track of products they are interested in without purchasing immediately. By making it easier to return to those items later, it increases the likelihood that consumers eventually complete the purchase.

**Browsing frequency (+):**  
Browsing frequency is expected to have a positive effect on purchase frequency. Consumers who browse Amazon more often are exposed to more products, recommendations, and promotions. Greater exposure increases the chances that they will find something they want to buy.

**Product category effects (+ / - depending on category):**  
Product category may also affect purchase frequency. Categories such as groceries and clothing are usually purchased more regularly, so they are expected to be positively associated with purchase frequency. In contrast, categories such as electronics are often bought less frequently because they are more expensive and more durable, so their effect may be smaller or negative relative to frequently purchased goods.

**Summary:**  
Overall, variables that reduce uncertainty, lower search costs, or increase consumer engagement are expected to have a positive relationship with purchase frequency on Amazon. 

# 2. Variable Selection 

## (a) Boruta Algorithm (Quantitative Predictors)

To identify the most important quantitative predictors of purchase frequency, we use the Boruta feature selection algorithm. Boruta is a wrapper method based on random forests that evaluates the importance of each variable by comparing it to randomized versions of the same variables. Variables that consistently perform better than their randomized counterparts are considered important predictors.

For this analysis, we used the Boruta package with the default random forest parameters and 100 maximum runs.

The Boruta algorithm identified four confirmed important predictors of purchase frequency: customer_reviews_importance, personalized_recommendation_frequency, rating_accuracy, and shopping_satisfaction. One additional variable, age, was classified as tentative but still potentially relevant. This analysis examines several variables that may explain differences in purchase frequency, and five quantitative predictors were selected for the final analysis.

```{r}
# Convert purchase frequency to numeric for Boruta
df$purchase_frequency_num <- as.numeric(as.factor(df$purchase_frequency))

# Keep only numeric variables
numeric_vars <- df %>%
  select(where(is.numeric))

# Run Boruta
boruta_model <- Boruta(purchase_frequency_num ~ ., data = numeric_vars, maxRuns = 100)

# View results
print(boruta_model)

# Plot variable importance
plot(boruta_model)
```

To identify relevant categorical predictors, we conducted ANOVA tests to examine whether differences in purchase frequency were associated with gender, purchase categories, and browsing frequency. The results indicate that gender and purchase categories do not significantly affect purchase frequency at conventional significance levels. Browsing frequency shows the strongest relationship with purchase frequency (p ≈ 0.092), suggesting that consumers who browse the platform more frequently may be more likely to make purchases. Based on these exploratory results, browsing frequency and purchase categories were selected as factor variables for the remaining analysis.

```{r}
# Convert categorical variables to factors
df$gender <- as.factor(df$gender)
df$purchase_categories <- as.factor(df$purchase_categories)
df$browsing_frequency <- as.factor(df$browsing_frequency)

# ANOVA tests
anova_gender <- aov(purchase_frequency_num ~ gender, data = df)
summary(anova_gender)

anova_category <- aov(purchase_frequency_num ~ purchase_categories, data = df)
summary(anova_category)

anova_browsing <- aov(purchase_frequency_num ~ browsing_frequency, data = df)
summary(anova_browsing)
```

## (c) Final Predictor Set 
Based on the Boruta algorithm and exploratory analysis of categorical variables, the final predictors used in the remainder of the analysis include the quantitative variables customer_reviews_importance, personalized_recommendation_frequency, rating_accuracy, shopping_satisfaction, and age. In addition, the factor variables purchase_categories and browsing_frequency are included to capture differences in consumer purchasing behavior across product types and browsing habits. 

# 3. Descripitve Analysis and Exploratory Data Analysis 
## (a) Dataset Description 
Describe: dataset source, time period, sample, variable definitions

## (b) Exploratory Visualizations 
# histograms
# scatterplots
# correlation plots

## (c) Descripitve Statistics Table 
# summary statistics 

## (d) Density Plots 
# density plots 

## (e) Nonlinearity and Transformations 
Discuss transformations such as log or polynomial terms 

## (f) Outliers 
Explain any outliers and how you handled them 

## (g) Missing Data 
Explain NA removal or imputation 

# 4. Model Building and Selection 
Model 1 
#regression model 

Model 2 
#regreession model

Model 3 
#regreession model 

Multicollinearity (VIF)
#VIF test 

Model Misspecification (RESET test)
#RESET test 

Diagnostic Plots 
#residual plots 

Model Comparison 
# AIC / BIC
Explain:

#5. Model Diagnostics and Robustness 
Linearity: Residual v. Fitted Plots 

Homoskedasticity 
#Breusch-Pagan test 

Normality 
# QQ plot
# Jarque-Bera test

Outliers and Influence
# Cook's distance

Bootstrap Analysis
# bootstrap code

Cross Validation
# 5-fold or 10-fold CV

# 6. Results and Economic Interpretation
# regression table
Explain: coefficient signs,economic meaning, statistical significance, confidence intervals, R²

# 7. Limitations and Extensions
data limitations, potential bias, possible future research

# 8. Code Quality and Reproducibility
Explain: reproducibility, seed setting, well-commented code

